---
title: "03_01_Explore_Enviro"
output: html_notebook
---

#Environmental Data Stats

Set-up workspace
```{r}
library(sp)
library(gstat)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(rcompanion)

#Read in environmental data
data <- read.csv("./soils_data_w2_parse.csv")

#Save the summary data as a csv file in output
sum <- summary(data)
write.csv(sum, "./output/enviro_summary_data.csv")

#Remove all data we dont need (like cluster, name, elements below detection levels)
vars <- c("lon","lat","pH","EC","GWC","per_clay","per_sand","per_silt",
          "Al","Ca","Co","Cr","Cu","Fe","K","Mg","Mn","Na","Ni","P","S",
          "Ti","Zn","per_som","PerN","PerC","CNRatio","Ele")
sd.full <- data[,vars]

#Create a SPDF class object so that the data is clearly distinguished from the coordinates
coordinates(sd.full) <- ~ lon + lat
```


#Check the distribution of each variable - looking for normality

There are several variables that have non-normal distributions. The first step to resolve this before applying a variogram is to normalize our data. We can start by log transforming the ones that appear non-normal and re-running our normality checks. 


```{r}

#For each soil property, run this program, changing the names accordingly
var.test = sd.full$per_sand_log
var.name = "Log of % Sand"

plot(var.test,
     main = var.name,
     xlab = "Sample Number")
hist(var.test,
     main = paste("Histogram of", var.name),
     breaks = 50)
ggdensity(var.test,
          main = paste("Density of", var.name),
          xlab = var.name)
ggqqplot(var.test)
shapiro.test(var.test)

##Check pH
#pH has a normal distribution

##Check EC
#EC has a non-normal distribution, try log transformation
sd.full$EC_log <- log(sd.full$EC)
#log transformation does not seem to fix this... rather, now we have  a very strong outlier

##Check GWC
#GWC has a non-normal distribution
sd.full$GWC_log <- log(sd.full$GWC)
#log transformation does not result in normality

##Check per_clay
#Clay has a non-normal distribution
sd.full$per_clay_log <- log(sd.full$per_clay)
#log transformation does not result in normality, data is still left-skewed

##Check per_sand
#Sand has a non-normal distribution
sd.full$per_sand_log <- log(sd.full$per_sand)
#log transformation results IN NORMALITY

##Check per_silt
#Silt has a non-normal distribution
sd.full$per_silt_log <- log(sd.full$per_silt)
#log transformation does not result in normality, data is left-skewed, one possible outlier

##Check Al
#Al has a normal distribution

##Check Ca
#Ca has a normal distribution

##Check Co
#Co has a non-normal distribution
sd.full$Co_log <- log(sd.full$Co)
#log transformation results IN NORMALITY (but the QQ plot looks a little odd)

##Check Cr
#Cr has a non-normal distribution
sd.full$Cr_log <- log(sd.full$Cr)
#log transformation does not result in normality, slightly right skewed

##Check Cu
#Cu has a non-normal distribution
sd.full$Cu_log <- log(sd.full$Cu)
#log transformation does not result in normality, left skewed

##Check Fe
#Fe has a non-normal distribution
sd.full$Fe_log <- log(sd.full$Fe)
#log transformation results IN NORMALITY

##Check K
#K has a non-normal distribution and one possible outlier
sd.full$K_log <- log(sd.full$K)
#log transformation does not result in normality... VERY suspicous looking outlier(s?)

##Check Mg
#Mg has a non-normal distribution
sd.full$Mg_log <- log(sd.full$Mg)
#log transformation does not result in normality, very left skewed

##Check Mn
#Mn has a non-normal distribution
sd.full$Mn_log <- log(sd.full$Mn)
#log transformation results IN NORMALITY

##Check Na
#Na has a non-normal distribution
sd.full$Na_log <- log(sd.full$Na)
#log transformation does not result in normality p=0.02, slightly left skewed

##Check Ni
#Ni has a non-normal distribution
sd.full$Ni_log <- log(sd.full$Ni)
#log transformation does not result in normality, right skewed/almost bimodal

##Check P
#P has a non-normal distribution
sd.full$P_log <- log(sd.full$P)
#log transformation does not result in normality, right skewed/almost bimodal

##Check S
#S has a non-normal distribution
sd.full$S_log <- log(sd.full$S)
#log transformation does not result in normality, very right skewed/almost bimodal

##Check Ti
#Ti has a non-normal distribution, possible outlier
sd.full$Ti_log <- log(sd.full$Ti)
#log transformation does not result in normality, right skewed

##Check Zn
#Zn has a non-normal distribution
sd.full$Zn_log <- log(sd.full$Zn)
#log transformation does not result in normality, ONE EXTREMELY SUSPECT OUTLIER

##Check per_som
#per_som has a normal distribution

##Check PerN
#PerN has a non-normal distribution, one data point looks slightly suspect (outlier)
sd.full$PerN_log <- log(sd.full$PerN)
#log transformation does not result in normality, ONE EXTREMELY SUSPECT OUTLIER

##Check PerC
#PerC has a non-normal distribution, one data point looks slightly suspect (outlier)
sd.full$PerC_log <- log(sd.full$PerC)
#log transformation does not result in normality, ONE EXTREMELY SUSPECT OUTLIER

##Check CNRatio
#CNRatio has a normal distribution

```

There are several variables that have non-normal distributions. The first step to resolve this before applying a variogram is to normalize our data. We can start by log transforming the ones that appear non-normal and re-running our normality checks. 









#Create variogram models for each variable

#There are several variogram models to choose from
#vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc", "Ste", "Cir", "Lin", "Bes", "Pen", "Per", "Wav", "Hol", "Log", "Pow", "Spl")))



lzn.vgm.ph <- variogram((pH)~lon+lat, sd.full)
lzn.fit.ph <- fit.variogram(lzn.vgm.ph, model=
                              vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.EC <- variogram((EC)~lon+lat, sd.full)
lzn.fit.EC <- fit.variogram(lzn.vgm.EC, model=
                              vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.GWC <- variogram((GWC)~lon+lat, sd.full)
lzn.fit.GWC <- fit.variogram(lzn.vgm.GWC, model=
                               vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.per_clay <- variogram((per_clay)~lon+lat, sd.full)
lzn.fit.per_clay <- fit.variogram(lzn.vgm.per_clay, model=
                                    vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.per_sand <- variogram((per_sand)~lon+lat, sd.full)
lzn.fit.per_sand <- fit.variogram(lzn.vgm.per_sand, model=
                                    vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.per_silt <- variogram((per_silt)~lon+lat, sd.full)
lzn.fit.per_silt <- fit.variogram(lzn.vgm.per_silt, model=
                                    vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.Al <- variogram((Al)~lon+lat, sd.full)
lzn.fit.Al <- fit.variogram(lzn.vgm.Al, model=
                              vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.Ca <- variogram((Ca)~lon+lat, sd.full)
lzn.fit.Ca <- fit.variogram(lzn.vgm.Ca, model=
                              vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.Co <- variogram((Co)~lon+lat, sd.full)
lzn.fit.Co <- fit.variogram(lzn.vgm.Co, model=
                              vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.Cr <- variogram((Cr)~lon+lat, sd.full)
lzn.fit.Cr <- fit.variogram(lzn.vgm.Cr, model=
                              vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.Cu <- variogram((Cu)~lon+lat, sd.full)
lzn.fit.Cu <- fit.variogram(lzn.vgm.Cu, model=
                              vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.Fe <- variogram((Fe)~lon+lat, sd.full)
lzn.fit.Fe <- fit.variogram(lzn.vgm.Fe, model=
                              vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.K <- variogram((K)~lon+lat, sd.full)
lzn.fit.K <- fit.variogram(lzn.vgm.K, model=
                             vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.Mg <- variogram((Mg)~lon+lat, sd.full)
lzn.fit.Mg <- fit.variogram(lzn.vgm.Mg, model=
                              vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.Mn <- variogram((Mn)~lon+lat, sd.full)
lzn.fit.Mn <- fit.variogram(lzn.vgm.Mn, model=
                              vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.Na <- variogram((Na)~lon+lat, sd.full)
lzn.fit.Na <- fit.variogram(lzn.vgm.Na, model=
                              vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.Ni <- variogram((Ni)~lon+lat, sd.full)
lzn.fit.Ni <- fit.variogram(lzn.vgm.Ni, model=
                              vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.P <- variogram((P)~lon+lat, sd.full)
lzn.fit.P <- fit.variogram(lzn.vgm.P, model=
                             vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.S <- variogram((S)~lon+lat, sd.full)
lzn.fit.S <- fit.variogram(lzn.vgm.S, model=
                             vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.Ti <- variogram((Ti)~lon+lat, sd.full)
lzn.fit.Ti <- fit.variogram(lzn.vgm.Ti, model=
                              vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.Zn <- variogram((Zn)~lon+lat, sd.full)
lzn.fit.Zn <- fit.variogram(lzn.vgm.Zn, model=
                              vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.per_som <- variogram((per_som)~lon+lat, sd.full)
lzn.fit.per_som <- fit.variogram(lzn.vgm.per_som, model=
                                   vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.PerN <- variogram((PerN)~lon+lat, sd.full)
lzn.fit.PerN <- fit.variogram(lzn.vgm.PerN, model=
                                vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.PerC <- variogram((PerC)~lon+lat, sd.full)
lzn.fit.PerC <- fit.variogram(lzn.vgm.PerC, model=
                                vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.CNRatio <- variogram((CNRatio)~lon+lat, sd.full)
lzn.fit.CNRatio <- fit.variogram(lzn.vgm.CNRatio, model=
                                   vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.Ele <- variogram((Ele)~lon+lat, sd.full)
lzn.fit.Ele <- fit.variogram(lzn.vgm.Ele, model=
                               vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))


#Plot variograms
plot(main="pH", lzn.vgm.ph, lzn.fit.ph)
plot(main="EC", lzn.vgm.EC, lzn.fit.EC)
plot(main="GWC", lzn.vgm.GWC, lzn.fit.GWC)
plot(main="% Clay", lzn.vgm.per_clay, lzn.fit.per_clay)
plot(main="% Silt", lzn.vgm.per_silt, lzn.fit.per_silt)
plot(main="% Sand", lzn.vgm.per_sand, lzn.fit.per_sand)
plot(main="Al", lzn.vgm.Al, lzn.fit.Al)
plot(main="Ca", lzn.vgm.Ca, lzn.fit.Ca)
plot(main="Co", lzn.vgm.Co, lzn.fit.Co)
plot(main="Cr", lzn.vgm.Cr, lzn.fit.Cr)
plot(main="Cu", lzn.vgm.Cu, lzn.fit.Cu)
plot(main="Fe", lzn.vgm.Fe, lzn.fit.Fe)
plot(main="K", lzn.vgm.K, lzn.fit.K)
plot(main="Mg", lzn.vgm.Mg, lzn.fit.Mg)
plot(main="Mn", lzn.vgm.Mn, lzn.fit.Mn)
plot(main="Na", lzn.vgm.Na, lzn.fit.Na)
plot(main="Ni", lzn.vgm.Ni, lzn.fit.Ni)
plot(main="P", lzn.vgm.P, lzn.fit.P)
plot(main="S", lzn.vgm.S, lzn.fit.S)
plot(main="Ti", lzn.vgm.Ti, lzn.fit.Ti)
plot(main="Zn", lzn.vgm.Zn, lzn.fit.Zn)
plot(main="per_som", lzn.vgm.per_som, lzn.fit.per_som)
plot(main="PerN", lzn.vgm.PerN, lzn.fit.PerN)
plot(main="PerC", lzn.vgm.PerC, lzn.fit.PerC)
plot(main="CNRatio", lzn.vgm.CNRatio, lzn.fit.CNRatio)
plot(main="Ele", lzn.vgm.Ele, lzn.fit.Ele)

#Create grids from the variograms to perfrom ordinary kriging - this will help us map the variables in QGIS
##Determine range of lat and lon
bbox(sd.full)

##Create a grid to esimate values over
x_range <- as.numeric(c(-94.23546, -94.23402))
y_range <- as.numeric(c(36.06563, 36.06665))
# create an empty grid of values ranging from the xmin-xmax, ymin-ymax
sample.grid <- expand.grid(x = seq(from = x_range[1],
                                   to = x_range[2], 
                                   length.out=30),
                           y = seq(from = y_range[1],                                           to = y_range[2], 
                                   length.out=30))  # expand points to grid
class(sample.grid)

sample.grid$number=seq(from=1, to=900, length.out=900)
plot1 <- sd.full %>% as.data.frame %>%
  ggplot(aes(lat, lon)) + geom_point(size=.25) + coord_equal() + 
  ggtitle("Points with measurements")
# this is clearly gridded over the region of interest
plot2 <- sample.grid %>% as.data.frame %>%
  ggplot(aes(y, x)) + geom_point(size=.25) + coord_equal() + 
  ggtitle("Points at which to estimate")
library(gridExtra)
grid.arrange(plot1, plot2, ncol=2)

coordinates(sample.grid) <- ~ x + y

lzn.kriged.ph <- krige(ph~1, sd.full, sample.grid, model=lzn.fit.ph)
