---
title: "03_01_Explore_Enviro"
output: html_notebook
---

#Environmental Data Stats

Set-up workspace
```{r}
library(sp)
library(gstat)
library(dplyr)
library(ggplot2)
library(ggpubr)

#Read in environmental data
data <- read.csv("./soils_data_w2_parse.csv")

#Save the summary data as a csv file in output
sum <- summary(data)
write.csv(sum, "./output/enviro_summary_data.csv")

#Remove all data we dont need (like cluster, name, elements below detection levels)
vars <- c("lon","lat","pH","EC","GWC","per_clay","per_sand","per_silt",
          "Al","Ca","Co","Cr","Cu","Fe","K","Mg","Mn","Na","Ni","P","S",
          "Ti","Zn","per_som","PerN","PerC","CNRatio","Ele")
sd.full <- data[,vars]

#Create a SPDF class object so that the data is clearly distinguished from the coordinates
coordinates(sd.full) <- ~ lon + lat
```


#Check the distribution of each variable - looking for normality

```{r}
##Check pH
plot(sd.full$pH)
ggdensity(sd.full$pH,
          main = "Density of pH",
          xlab = "pH")
ggqqplot(sd.full$pH)
shapiro.test(sd.full$pH)
#pH has a normal distribution

##Check EC
plot(sd.full$EC)
ggdensity(sd.full$EC,
          main = "Density of EC",
          xlab = "EC")
ggqqplot(sd.full$EC)
shapiro.test(sd.full$EC)
#EC has a non-normal distribution 

##Check GWC
plot(sd.full$GWC)
ggdensity(sd.full$GWC,
          main = "Density of GWC",
          xlab = "GWC")
ggqqplot(sd.full$GWC)
shapiro.test(sd.full$GWC)
#GWC has a non-normal distribution

##Check per_clay
plot(sd.full$per_clay)
ggdensity(sd.full$per_clay,
          main = "Density of per_clay",
          xlab = "per_clay")
ggqqplot(sd.full$per_clay)
shapiro.test(sd.full$per_clay)
#Clay has a non-normal distribution

##Check per_sand
plot(sd.full$per_sand)
ggdensity(sd.full$per_sand,
          main = "Density of per_sand",
          xlab = "per_sand")
ggqqplot(sd.full$per_sand)
shapiro.test(sd.full$per_sand)
#Sand has a non-normal distribution

##Check per_silt
plot(sd.full$per_silt)
ggdensity(sd.full$per_silt,
          main = "Density of per_silt",
          xlab = "per_silt")
ggqqplot(sd.full$per_silt)
shapiro.test(sd.full$per_silt)
#Silt is non-normal

##Check Al
ggdensity(sd.full$Al,
          main = "Density of Al",
          xlab = "Al")
ggqqplot(sd.full$Al)
shapiro.test(sd.full$Al)
#Al has a normal distribution

##Check Ca
ggdensity(sd.full$Ca,
          main = "Density of Ca",
          xlab = "Ca")
ggqqplot(sd.full$Ca)
shapiro.test(sd.full$Ca)
#Ca has a normal distribution

##Check Co
ggdensity(sd.full$Co,
          main = "Density of Co",
          xlab = "Co")
ggqqplot(sd.full$Co)
shapiro.test(sd.full$Co)
#Co has a non-normal distribution

##Check Cr
ggdensity(sd.full$Cr,
          main = "Density of Cr",
          xlab = "Cr")
ggqqplot(sd.full$Cr,
         main = "qq plot for Cr")
shapiro.test(sd.full$Cr)
#Cr has a non-normal distribution

##Check Cu
ggdensity(sd.full$Cu,
          main = "Density of Cu",
          xlab = "Cu")
ggqqplot(sd.full$Cu,
         main = "qq plot for Cu")
shapiro.test(sd.full$Cu)
#Cu has a non-normal distribution

##Check Fe
ggdensity(sd.full$Fe,
          main = "Density of Fe",
          xlab = "Fe")
ggqqplot(sd.full$Cu,
         main = "qq plot for Fe")
shapiro.test(sd.full$Fe)
#Fe has a non-normal distribution

##Check K
ggdensity(sd.full$K,
          main = "Density of K",
          xlab = "K")
ggqqplot(sd.full$K,
         main = "qq plot for K")
shapiro.test(sd.full$K)
#K has a non-normal distribution and one possible outlier

##Check Mg
ggdensity(sd.full$Mg,
          main = "Density of Mg",
          xlab = "Mg")
ggqqplot(sd.full$Mg,
         main = "qq plot for Mg")
shapiro.test(sd.full$Mg)
#Mg has a non-normal distribution

##Check Mn
ggdensity(sd.full$Mn,
          main = "Density of Mn",
          xlab = "Mn")
ggqqplot(sd.full$Mn,
         main = "qq plot for Mn")
shapiro.test(sd.full$Mn)
#Mn has a non-normal distribution

##Check Na
ggdensity(sd.full$Na,
          main = "Density of Na",
          xlab = "Na")
ggqqplot(sd.full$Na,
         main = "qq plot for Na")
shapiro.test(sd.full$Na)
#Na has a non-normal distribution

##Check Ni
ggdensity(sd.full$Ni,
          main = "Density of Ni",
          xlab = "Ni")
ggqqplot(sd.full$Ni,
         main = "qq plot for Ni")
shapiro.test(sd.full$Ni)
#Ni has a non-normal distribution

##Check P
ggdensity(sd.full$P,
          main = "Density of P",
          xlab = "P")
ggqqplot(sd.full$P,
         main = "qq plot for P")
shapiro.test(sd.full$P)
#P has a non-normal distribution

##Check S
ggdensity(sd.full$S,
          main = "Density of S",
          xlab = "S")
ggqqplot(sd.full$S,
         main = "qq plot for S")
shapiro.test(sd.full$S)
#S has a non-normal distribution

##Check Ti
plot(sd.full$Ti)
ggdensity(sd.full$Ti,
          main = "Density of Ti",
          xlab = "Ti")
ggqqplot(sd.full$Ti,
         main = "qq plot for Ti")
shapiro.test(sd.full$Ti)
#Ti has a non-normal distribution, possible outlier

##Check Zn
plot(sd.full$Zn)
ggdensity(sd.full$Zn,
          main = "Density of Zn",
          xlab = "Zn")
ggqqplot(sd.full$Zn,
         main = "qq plot for Zn")
shapiro.test(sd.full$Zn)
#Zn has a non-normal distribution


##Check per_som
plot(sd.full$per_som)
ggdensity(sd.full$per_som,
          main = "Density of per_som",
          xlab = "per_som")
ggqqplot(sd.full$per_som,
         main = "qq plot for per_som")
shapiro.test(sd.full$per_som)
#per_som has a non-normal distribution, two possible outliers


##Check PerN
plot(sd.full$PerN)
ggdensity(sd.full$PerN,
          main = "Density of PerN",
          xlab = "PerN")
ggqqplot(sd.full$PerN,
         main = "qq plot for PerN")
shapiro.test(sd.full$PerN)
#PerN has a non-normal distribution, one data point looks slightly suspect (outlier)


##Check PerN
plot(sd.full$PerC)
ggdensity(sd.full$PerC,
          main = "Density of PerC",
          xlab = "PerC")
ggqqplot(sd.full$PerC,
         main = "qq plot for PerC")
shapiro.test(sd.full$PerC)
#PerC has a non-normal distribution, one data point looks slightly suspect (outlier)


##Check CNRatio
plot(sd.full$CNRatio)
ggdensity(sd.full$CNRatio,
          main = "Density of CNRatio",
          xlab = "CNRatio")
ggqqplot(sd.full$PerN,
         main = "qq plot for CNRatio")
shapiro.test(sd.full$CNRatio)
#CNRatio has a normal distribution

##Check Ele
plot(sd.full$Ele)
ggdensity(sd.full$Ele,
          main = "Density of Ele",
          xlab = "Ele")
ggqqplot(sd.full$Ele,
         main = "qq plot for Ele")
shapiro.test(sd.full$Ele)
#Elevation has a non-normal distribution





```



#Create variogram models for each variable

#There are several variogram models to choose from
#vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc", "Ste", "Cir", "Lin", "Bes", "Pen", "Per", "Wav", "Hol", "Log", "Pow", "Spl")))



lzn.vgm.ph <- variogram((pH)~lon+lat, sd.full)
lzn.fit.ph <- fit.variogram(lzn.vgm.ph, model=
                              vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.EC <- variogram((EC)~lon+lat, sd.full)
lzn.fit.EC <- fit.variogram(lzn.vgm.EC, model=
                              vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.GWC <- variogram((GWC)~lon+lat, sd.full)
lzn.fit.GWC <- fit.variogram(lzn.vgm.GWC, model=
                               vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.per_clay <- variogram((per_clay)~lon+lat, sd.full)
lzn.fit.per_clay <- fit.variogram(lzn.vgm.per_clay, model=
                                    vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.per_sand <- variogram((per_sand)~lon+lat, sd.full)
lzn.fit.per_sand <- fit.variogram(lzn.vgm.per_sand, model=
                                    vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.per_silt <- variogram((per_silt)~lon+lat, sd.full)
lzn.fit.per_silt <- fit.variogram(lzn.vgm.per_silt, model=
                                    vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.Al <- variogram((Al)~lon+lat, sd.full)
lzn.fit.Al <- fit.variogram(lzn.vgm.Al, model=
                              vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.Ca <- variogram((Ca)~lon+lat, sd.full)
lzn.fit.Ca <- fit.variogram(lzn.vgm.Ca, model=
                              vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.Co <- variogram((Co)~lon+lat, sd.full)
lzn.fit.Co <- fit.variogram(lzn.vgm.Co, model=
                              vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.Cr <- variogram((Cr)~lon+lat, sd.full)
lzn.fit.Cr <- fit.variogram(lzn.vgm.Cr, model=
                              vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.Cu <- variogram((Cu)~lon+lat, sd.full)
lzn.fit.Cu <- fit.variogram(lzn.vgm.Cu, model=
                              vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.Fe <- variogram((Fe)~lon+lat, sd.full)
lzn.fit.Fe <- fit.variogram(lzn.vgm.Fe, model=
                              vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.K <- variogram((K)~lon+lat, sd.full)
lzn.fit.K <- fit.variogram(lzn.vgm.K, model=
                             vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.Mg <- variogram((Mg)~lon+lat, sd.full)
lzn.fit.Mg <- fit.variogram(lzn.vgm.Mg, model=
                              vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.Mn <- variogram((Mn)~lon+lat, sd.full)
lzn.fit.Mn <- fit.variogram(lzn.vgm.Mn, model=
                              vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.Na <- variogram((Na)~lon+lat, sd.full)
lzn.fit.Na <- fit.variogram(lzn.vgm.Na, model=
                              vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.Ni <- variogram((Ni)~lon+lat, sd.full)
lzn.fit.Ni <- fit.variogram(lzn.vgm.Ni, model=
                              vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.P <- variogram((P)~lon+lat, sd.full)
lzn.fit.P <- fit.variogram(lzn.vgm.P, model=
                             vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.S <- variogram((S)~lon+lat, sd.full)
lzn.fit.S <- fit.variogram(lzn.vgm.S, model=
                             vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.Ti <- variogram((Ti)~lon+lat, sd.full)
lzn.fit.Ti <- fit.variogram(lzn.vgm.Ti, model=
                              vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.Zn <- variogram((Zn)~lon+lat, sd.full)
lzn.fit.Zn <- fit.variogram(lzn.vgm.Zn, model=
                              vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.per_som <- variogram((per_som)~lon+lat, sd.full)
lzn.fit.per_som <- fit.variogram(lzn.vgm.per_som, model=
                                   vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.PerN <- variogram((PerN)~lon+lat, sd.full)
lzn.fit.PerN <- fit.variogram(lzn.vgm.PerN, model=
                                vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.PerC <- variogram((PerC)~lon+lat, sd.full)
lzn.fit.PerC <- fit.variogram(lzn.vgm.PerC, model=
                                vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.CNRatio <- variogram((CNRatio)~lon+lat, sd.full)
lzn.fit.CNRatio <- fit.variogram(lzn.vgm.CNRatio, model=
                                   vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))

lzn.vgm.Ele <- variogram((Ele)~lon+lat, sd.full)
lzn.fit.Ele <- fit.variogram(lzn.vgm.Ele, model=
                               vgm(c("Exp","Sph", "Gau", "Mat", "Nug", "Exc")))


#Plot variograms
plot(main="pH", lzn.vgm.ph, lzn.fit.ph)
plot(main="EC", lzn.vgm.EC, lzn.fit.EC)
plot(main="GWC", lzn.vgm.GWC, lzn.fit.GWC)
plot(main="% Clay", lzn.vgm.per_clay, lzn.fit.per_clay)
plot(main="% Silt", lzn.vgm.per_silt, lzn.fit.per_silt)
plot(main="% Sand", lzn.vgm.per_sand, lzn.fit.per_sand)
plot(main="Al", lzn.vgm.Al, lzn.fit.Al)
plot(main="Ca", lzn.vgm.Ca, lzn.fit.Ca)
plot(main="Co", lzn.vgm.Co, lzn.fit.Co)
plot(main="Cr", lzn.vgm.Cr, lzn.fit.Cr)
plot(main="Cu", lzn.vgm.Cu, lzn.fit.Cu)
plot(main="Fe", lzn.vgm.Fe, lzn.fit.Fe)
plot(main="K", lzn.vgm.K, lzn.fit.K)
plot(main="Mg", lzn.vgm.Mg, lzn.fit.Mg)
plot(main="Mn", lzn.vgm.Mn, lzn.fit.Mn)
plot(main="Na", lzn.vgm.Na, lzn.fit.Na)
plot(main="Ni", lzn.vgm.Ni, lzn.fit.Ni)
plot(main="P", lzn.vgm.P, lzn.fit.P)
plot(main="S", lzn.vgm.S, lzn.fit.S)
plot(main="Ti", lzn.vgm.Ti, lzn.fit.Ti)
plot(main="Zn", lzn.vgm.Zn, lzn.fit.Zn)
plot(main="per_som", lzn.vgm.per_som, lzn.fit.per_som)
plot(main="PerN", lzn.vgm.PerN, lzn.fit.PerN)
plot(main="PerC", lzn.vgm.PerC, lzn.fit.PerC)
plot(main="CNRatio", lzn.vgm.CNRatio, lzn.fit.CNRatio)
plot(main="Ele", lzn.vgm.Ele, lzn.fit.Ele)

#Create grids from the variograms to perfrom ordinary kriging - this will help us map the variables in QGIS
##Determine range of lat and lon
bbox(sd.full)

##Create a grid to esimate values over
x_range <- as.numeric(c(-94.23546, -94.23402))
y_range <- as.numeric(c(36.06563, 36.06665))
# create an empty grid of values ranging from the xmin-xmax, ymin-ymax
sample.grid <- expand.grid(x = seq(from = x_range[1],
                                   to = x_range[2], 
                                   length.out=30),
                           y = seq(from = y_range[1],                                           to = y_range[2], 
                                   length.out=30))  # expand points to grid
class(sample.grid)

sample.grid$number=seq(from=1, to=900, length.out=900)
plot1 <- sd.full %>% as.data.frame %>%
  ggplot(aes(lat, lon)) + geom_point(size=.25) + coord_equal() + 
  ggtitle("Points with measurements")
# this is clearly gridded over the region of interest
plot2 <- sample.grid %>% as.data.frame %>%
  ggplot(aes(y, x)) + geom_point(size=.25) + coord_equal() + 
  ggtitle("Points at which to estimate")
library(gridExtra)
grid.arrange(plot1, plot2, ncol=2)

coordinates(sample.grid) <- ~ x + y

lzn.kriged.ph <- krige(ph~1, sd.full, sample.grid, model=lzn.fit.ph)
